---
title: "Red wine quality EDA "
author: "Taef al khales"
output: 
  html_document:
      toc: true
      
knit: (function(input, ...) {rmarkdown::render(input,output_dir = "../")})

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library (skimr)
library(corrgram)
library(ggpubr)
library(reactable)
library(GGally)
library(plotly)
library(here)

#here 
here("individual-assignment-2-Taifabdullah2021","wine_report.rmd")

#reading the data
wine_og <- read_csv("../data/winequality-red.csv")

```


##  <span style="color:blue">1. The Data</span>

- *fixed acidity*: most acids involved with wine or fixed or nonvolatile.
- *volatile acidity*:the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste
- citric acid:found in small quantities, citric acid can add ‘freshness’ and flavor to wines
- *residual sugar*:  the amount of sugar remaining after fermentation stops, it’s rare to find wines with less than 1 gram/liter and wines with greater than 45 grams/liter are considered sweet
- *chlorides*: the amount of salt in the wine
- *free sulfur dioxide* : the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion
- *total sulfur dioxide*: amount of free and bound forms of S02; in low concentrations
- *density*: the density of water is close to that of water depending on the percent alcohol and sugar content
- *pH*:describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
- *sulphates*: a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant
- *alcohol*:the percent alcohol content of the wine
- *quality* (score between 0 and 10)


----------------------------------------------------


## 2. <span style="color:blue">EDA</span>

### 2.1 First look at the data 

```{r echo=FALSE}
knitr::kable(head(wine_og), col.names = gsub("[.]", " ", names(wine_og)))

#reactable(wine_og)
```

### 2.2 Tidy up the columns names 

Now the names are seprated with (_) instead of a space. and  be in lower case.

```{r echo=FALSE}
wine_clean <- wine_og %>%
  janitor::clean_names()

reactable(wine_clean)
```

### 2.3 Summary statistics 

Skim is a very nice package for summary statistics [Skim](https://docs.ropensci.org/skimr/reference/skim.html).

```{r}
skim(wine_clean)
```

We can see the following things:

1. all columns are numeric 
2. no missing values
3. it is not a big dataset



### 2.4 Univariate

### 2.4.1 Quality Distribution

It appears that is this sample dataset, the most frequent wine quality are 5 and 6. 

```{r echo=FALSE}
ggplot(wine_clean,aes(x=quality))+geom_bar(stat = "count",position = "dodge", fill = "Red")+
  scale_x_continuous(breaks = seq(3,8,1))+
  ggtitle("Distribution of Red Wine Quality Ratings")+
  theme_classic()
```


### 2.4.2 PH Distribution (boxplot)

```{r echo=FALSE}
dens_ph <- ggplot(wine_clean)+geom_density(aes(x =p_h), colour='red')
boxplot_ph <- ggplot(wine_clean)+geom_boxplot(aes(x =p_h))

ggarrange(dens_ph,boxplot_ph,  nrow = 2)

```



### 2.4.3 All data Distribution
```{r echo=FALSE}
wine_clean %>%
  gather() %>% 
  ggplot(aes(value,fill=key)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(bins=sqrt(nrow(wine_clean))) +
    theme(legend.position="none") 
```


### 2.4.2 Alcohol Distribution (boxplot)

```{r echo=FALSE}
dens_al <- ggplot(wine_clean)+geom_density(aes(x =alcohol), colour='red')
boxplot_al <- ggplot(wine_clean)+geom_boxplot(aes(x =alcohol))

ggarrange(dens_al,boxplot_al,  nrow = 2)
```

### 2.5 Bivariate 

### 2.5.1 GGPair graph (group by quality)

```{r}
wine_ggpair <- wine_clean
wine_ggpair$quality <- as.factor(wine_ggpair$quality)
ggpairs(wine_ggpair,columns = 1:6,aes(colour = quality, alpha = 0.4), upper=list(continuous = "blank"))

wine_ggpair2 <- wine_clean
wine_ggpair2$quality <- as.factor(wine_ggpair2$quality)
ggpairs(wine_ggpair2,columns = 7:11,aes(colour = quality, alpha = 0.4), upper=list(continuous = "blank"))

```

### 2.5.2 Correlation (heatmap)

- *Alcohol* and *Quality* are inversely highly correlated.
- *fixed acidity* is highly correlated with *Ensity* and *Citric acid*
- *total sulfur dioxide* is highly correlated with *free sulfur dioxide*

```{r echo=FALSE}

corrgram(wine_clean, order=TRUE,
         upper.panel=panel.cor, main="correlation heat map")
```

### 2.6 Mean values grouped by the quality 

```{r}
wine_df <- wine_clean %>% group_by(quality) %>% summarise(across(everything(), mean))
wine_df <- signif(wine_df, digits = 3)
reactable(wine_df)
```



### 2.7 2D interractive graph


```{r}
plot <- ggplot(data=wine_clean, aes(alcohol,fixed_acidity)) +
geom_point(aes(color=quality)) +
ggtitle("2D interractive graph")

ggplotly(plot)
```

### 3. Conclusion 

In this report i have done some EDA for the Red Wine Quality from Kaggle. And it shows that some of the features like Alcohol and fixed_acidity may be good to build a prediction models the quality of the wine. And based on the heat map there were some inverse correlation between the values.

